Deep Reinforcement Learning for Conservation Decisions

### 2021-05-06 manuscript/manuscript.Rmd
most recet changes

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the uncertain and changing environments considered by RL research have largely been drawn from examples in robotic movement and video games.
RL agents have recently achieved superhuman performance on classic Atari games [@DQN], while Deepmind’s RL-trained AlphaZero has defeated the world's grandmasters in Go, Chess, and Shogi [@alphazero].
Complex environmental problems share many similarities to these games: the need to plan many moves ahead, accept short-term costs to achieve long-term benefit, to account for uncertainty and respond with contingency to the unexpected.
Of course real ecosystems are also far more complex than a game of chess or MoonLander -- but can the same be said of our efforts at finding the best strategy?
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler models are frequently employed throughout ecology to guide understanding and response.
<!-- Kari: ick, jargon heavy.  Need to say: some existing methods, but limited to simple problems or simple solutions -->
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of models, the so called "curse of dimensionality" makes problems with too many possible states or actions intractable by conventional methods such as stochastic dynamic programming (SDP) and approximate dynamic programming (ADP) [@Polasky2011; @Marescot2013].
Neural-network-based ML techniques have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the combination of such approaches from RL trained using sufficient simulations of suitable ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

Any application of RL can be divided into two components: an *environment* and an *agent*.
An *environment* is typically a computer simulation, though it is certainly possible to learn from real world environments, such as teaching physical robots to walk and navigate obstacles.
An *agent* is typically a computer algorithm which takes sequential observations from the environment and returns proposed actions. RL agents typically use one or more neural networks to represent the relationship between observed states and suitable actions, and 'train' these networks through repeated interactions with the environment to maximize the culumulative reward.
<!-- Blah blah skip this?-->
It is important to recognize that these definitions are significantly more general than requirements typically imposed by classical methods.  
For instance, there is no a priori assumption that the environment is *fully observed*, or that the environment is *Markovian* (future states depend only on the current state), commonly assumed by classical methods [e.g. @Marescot2013]. 
Just as important is the high degree of interoperability provided by the RL framework.
As we shall see, an RL algorithm developed to play Atari games can instead be trained on an environment that simulates fish population dynamics without modification.
(Note that this does not refer to the *trained* agent).
This not guarantee that such an algorithm will be successful (even after extensive tuning and training in the fish population dynamics environment), but does open the door to more rapid exchange between fields.
In particular, the adoption of standard interfaces for RL environments [@brockman2016] as well as the implementation of cutting edge algorithms in flexible open source frameworks [@sb3; @tf-agents] means that a researcher can test a range of RL algorithms against a single environment, or a given algorithm across a range of environments with equal ease.

>>>>

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@lillicrap; @alphazero; @alphastar]
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop their strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler simulators are frequently employed throughout ecology to guide understanding and response.
<!-- Kari: ick, jargon heavy.  Need to say: some existing methods, but limited to simple problems or simple solutions -->
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods such as stochastic dynamic programming (SDP) and approximate dynamic programming (ADP) [@Polasky2011; @Marescot2013].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the application of RL methods to simulations of ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

Any application of RL can be divided into two components: an *environment* and an *agent*.
An *environment* is typically a computer simulation, though it is possible to use the real world as the RL environment [@robotics; @offlinesurvey].
Over the course of an *episode*, an *agent*, which is often a computer program, continuously interacts with the environment. 
At each time step, the agent performs an *action*, and as a result of this action, the environment will determine the next *state* and numerical *reward* that the agent will observe.
The goal of the agent is to learn how to maximize its expected cumulative reward.
The agent learns how to achieve this objective during a period called *training*.
In training, the agent *explores* the available states and actions. 
Once the agent comes across a highly rewarding sequence of states and actions, the agent will reinforce this behavior so that it is more likely for the agent to *exploit* the same high reward trajectory in future episodes.
Throughout this process, the agent's behavior is codified into what is called a *policy*, which is a mapping of states to actions.

To solve the RL problem, algorithms either take a *model-free* or *model-based* approach. 
The distinction is that *model-free* algorithms do not have access to a model of the environment; yet, *model-based* algorithms use a model of the environment to achieve the RL objective.
The trade-off between these approaches is that when it is possible to quickly learn a good model or the model is already known, model-based algorithms tend to require much less interaction with the environment to learn good-performing policies. 
Yet, oftentimes learning a model of the environment is very difficult, and in these cases, model-free algorithms tend to outperform.
Since most of the recent focus in RL has gone towards model-free algorithms, we will focus on model-free agents for the rest of the paper.

Amongst model-free algorithms,
